### 2020-06-12 chapters/sota.tex
Kifer et. al - A two window paradigm algorithm for change detection

Quoting the authors, histogramming univariate point outlier detection methods are \textit{"based on detecting the points whose removal from the univariate time series results in a histogram representation with lower error than the original"}.
%TODO: explain this better, paper not clear to me

@@

\subsubsection{A two window paradigm algorithm for change detection}
\subsubsection*{Method classification}
\subsubsection*{Time complexity}
\subsubsection*{Space complexity}
\subsubsection*{Applicability to our Hypothesis}

\iffalse
Kifer et al. 2 window paradigm \\
ref window -> some data\\
sliding window -> current data\\
apply distance metric (like jsd) to both windows if above threshold alert and refill reference (adapts to new distro)\\
input data: uni or multi variate\\
method nature: univariate (doenst check correlation)\\
outlier type: subsequence outlier (window)\\
\fi

>>>>

Quoting the authors, histogramming univariate point outlier detection methods are \textit{"based on detecting the points whose removal from the univariate time series results in a histogram representation with lower error than the original"}.

\textcolor{red}{TODO: explain this better, paper not clear to me}

@@

\subsubsection{A two window paradigm algorithm for change detection}
Kifer et al. propose that detecting change in data streams can be reduced to testing if two windows have different underlying distributions \cite{Kifer-Detecting-Change}. The proposed method uses two windows, a reference \textit{W\textsubscript{1}} and a sliding one \textit{W\textsubscript{2}}. Both windows are tuple-based, have size \textit{k} and the sliding window is a true sliding one (recall Section \ref{sec:windows} and the expression \ref{math:windows}). The reference window \textit{W\textsubscript{2}} works as a \textit{reference of normality} and contains the first \textit{k} points of the stream that occurred after the last detected change.

The algorithm begins by filling both windows with the first \textit{k} tuples. Then it slides window \textit{W\textsubscript{2}} by one tuple, since it is a true sliding window, for each incoming event from the data stream. In between slides, the algorithm checks for change by applying a distance function \textit{d} taking as input both the reference and the sliding windows, which measures the discrepancy between both. Whenever this distance is above a certain threshold $\alpha$, the algorithm raises an alert. After raising an alert, both windows are re-initialized with the next \textit{k} items.

Note that after a change is detected, the reference window is updated with the next \textit{k} tuples from the stream. As we have stated, we want a method that does not adapt to the data stream changes and instead keeps raising alerts until the sliding window data distribution returns to the one seen in the reference period. It is trivial to adapt this method to always use the same reference period. To do so, we simply keep the original reference window and don't update it after each alert.

This algorithm reduces change detection in data streams to testing whether two samples --- \textit{reference} and \textit{sliding} windows --- are generated by different distributions. The authors study methods for identifying differences in distribution between two samples and implement them as the \textit{d} function.

\textcolor{red}{\textbf{TODO: develop this idea, they talk about Kolmogorov–Smirnov and others, maybe relate to model monitoring paper, talk about JSD, Wasserstein, so on so forth}}

\subsubsection*{Method classification}
The method devised by Kifer et. al can take uni or multivariate input (if we use one pair of reference/target windows for each time-series) but performs a univariate analysis as it does not find correlations between time-dependent variables. This two-windowed method measures the difference between both windows, the reference and the sliding one. Furthermore, instead of detecting point outliers, the alerts produced indicate the entire window (or subsequence) shows anomalous behavior. Hence, according to the taxonomy presented in Figure \ref{fig:subsequence-outlier-taxonomy}, this method is a dissimilarity-based univariate subsequence outlier detection method.

\subsubsection*{Time complexity}
The authors show that it is possible to compute the distance between both windows in logarithmic time regarding window size \textit{k}. Because the entire sliding window is held in memory, a window update is tantamount to removing the head of the list (element eviction) or appending a new element to the tail (element insertion), which is done in constant time. Thus, for each new event, the method evicts the oldest element, inserts the new one and computes the distance between the reference and sliding windows. In big O notation, the largest and determining factor is the computation of the distance function, done in \textit{O(log(k))} time, with \textit{k} as the size of both windows.

\subsubsection*{Space complexity}
The method makes use of two \textit{k}-sized windows, thus storing \textit{2k} elements. Hence, memory consumption grows linearly with window size.

\subsubsection*{Applicability to our Hypothesis}
In this Thesis, we want to detect and alert data pattern shifts with a low memory footprint. However, the method proposed by Kifer et. al grows linearly with window size, making it unfit for large windows, our use-case.
